\documentclass{article}
\input{preamble}
\title{A Language for Enthymemes}
\author{Max Rapp}
\begin{document}

\maketitle

\section{Introduction}

The notion of enthymemes traces back to Aristotle's rhetoric and is usually associated with two distinct kinds of ampliative inference steps (though the correct interpretation is disputed, see \cite{sep-aristotle-rhetoric}):

\begin{enumerate}
\item Deductive Arguments with missing premises (type-I-enthymemes);
\item Deductive Arguments with defeasible assumptions or inference steps (type-II-enthymemes).
\end{enumerate}

From a constructivist perspective both types of enthymemes contain subarguments that fail to furnish witnesses.
In this view the difference between the two consists merely in the burden-of-proof: type I enthymemes commit the proponent of the argument to present a witness if questioned; type II enthymemes instead commit any challenger to present a refuting witness.

\begin{oldpart}{Initial Braindump}
  In the previous paper we defined terms and types of the \matilda calculus and used free variables to encode assumptions.
  Here we introduce meta-variables to properly encode incomplete proofs (arguments), $\eta$-equivalence for arguments and an $\eta$-long normal form that serves as the basis for \emph{grafting} operations.
  The meta-variables we introduce come in two flavors: obligations assume the onus, requiring successful supporters to be accepted; presumptions delegate the onus and are hence accepted unless they are successfully defeated by a counterargument.

  \ednote{MR: Restriction: study only forward debates, i.e. typed forward: debate -> argument -> debate |= graft y on x. Types backward: argument -> debate -> debate |= graft y on x is not allowed. This ensures uniqueness of debate construction (if you graft a debate onto a debate, you would have to ensure that the scion arguments of the scion debate are universally grafted onto the rootstock debate, complicating the situation). This is not enough: the scion arguments' subarguments would still have to be grafted onto other open assumptions.}

  \ednote{Subtyping hierarchy: debate >= argument >= intuitionistic_proof}
  
\end{oldpart}

\section{Recap of the Term Language}

\begin{oldpart}{Missing}
  Here Recap of the previous paper
\end{oldpart}

We use catch, throw and steal patterns to encode support and attack between arguments:



  \begin{definition}[Attack]
    Let $t_1,t_2$ be intuitionistic \matilda expressions. An \emph{attack} between $t_1$ and $t_2$ is defined as one of the patterns in Grammar \ref{gram:attacks}.

    \begin{definition}[Support]
      Support is defined via the following grammar:
  
     \begin{grammar}[Support][][gram:support]
       \firstcasesubtil{$t_1\prec_\phi t_2$}{\catcht{: \alpha}.\rangle t_1 :\phi: \rangle \throwt{x }{:\alpha} :: \stealt{t_2}{:\alpha}\langle.\catchc{x:}\langle}{}
       \otherform{\rangle \catcht{:\alpha}.\rangle \stealc{\alpha}{x:}:: \throwc{t_2}{x:} \langle :\phi: t_2 \langle . \catchc{x:}}{}
     \end{grammar}
     \end{definition}


    \begin{grammar}[Attacks][][gram:attacks]
      \firstcasesubtil{$t_1 \leftattack_\phi t_2$}{\catcht{:\alpha}.\rangle t_1:\phi: \rangle \stealc{x:}{t_2} :: \throwc{x:}{\alpha}\langle.\catchc{x:} \langle}{}
      \otherform{\rangle \catcht{:\alpha}.\rangle \throwt{x}{:\alpha} :\phi: \stealt{t_2}{:\alpha}\langle:\phi: t_1\langle.\catchc{x:}}{}
    \end{grammar}

    We say that $t_2$ attacks $t_1$ on $\phi$ iff $t_1\leftattack_\phi t_2$. In particular, if $t_1$ is an assumption, $t_2$ \emph{undermines} $t_1$. Else $t_2$ \emph{rebuts} $t_1$
  \end{definition}

\section{Constructing Arguments: the Argument -calculus}

The \matilda calculus provides us with two of the three requirements we identified in \cref{sec:argum-debat-prop}: conflicting proof terms and control operators. The third feature - defeasibility - requires an extension of the language.

\subsection{Unwitnessed Premises: Obligations and Presumptions}


To deal with both kinds of enthymemes we extend the grammar of \matilda terms and environments by two types of \emph{unwitnessed premises}: \emph{obligations} and \emph{presumptions}:

\begin{grammar}[Argument Calculus grammar][tbhp][gram:terms]
  \firstcasesubtil{v}{\dots \gralt [:\sigma] \gralt \underline{[:\sigma]} }{Term obligations and presumptions}
  \firstcasesubtil{E}{\dots \gralt [\sigma:]\gralt \underline{[\sigma:]}}{Environment obligations and presumptions}
\end{grammar}

\begin{notation}\label[notation]{not:enthymeme-nominals}
  We overload notation and use $[:\sigma]$ for the typing judgment $[:\sigma]:\sigma$ and $[\sigma:]$ for the typing judgment $\sigma:[\sigma:]$. In cases where typing information is irrelevant we write $[.:]$ for term side obligations and $[:.]$ for environment side obligations. We use these notations for presumptions as well, with the obvious substitutions.
\end{notation}


Obligations are typed as follows:

\begin{center}
  \begin{minipage}{0.48\linewidth}
    \begin{prooftree}
      \label{rule:termenth}
      \infer0{\Sigma\vdash [:\sigma] \dashv \Delta}
    \end{prooftree}
  \end{minipage}

  \begin{minipage}{0.48\linewidth}
    \begin{prooftree}
      \label{rule:contextenth}
      \infer0{\Sigma\vdash [\sigma:] \dashv \Delta}
    \end{prooftree}
  \end{minipage}
\end{center}

Note that obligations are trivializing: one can find an argument for any proposition so long as one retains the obligations to prove one's assumptions.
On the other hand, presumptions do not have introduction rules: what presumptions are permitted in a given situation world knowledge. Examples of such knowledge are general burden-of-proof delegations such as negation-by-failure, or the assertion of looping proof obligations in coinductive inference as well as concrete default knowledge such as ``birds can usually fly''.\ednote{MR: work out these examples. E.g. coinductive inference is an axiom scheme like  forall x.P(x):- exists }
Thus reasoning with presumptions depends on the presense of a \emph{scenario}, i.e. some set of concrete presumption facts of the forms

\begin{center}
  \begin{minipage}{0.48\linewidth}
    \begin{prooftree}
      \label{rule:termenth}
      \infer0{\Sigma\vdash [:A] \dashv \Delta}
    \end{prooftree}
  \end{minipage}

  \begin{minipage}{0.48\linewidth}
    \begin{prooftree}
      \label{rule:contextenth}
      \infer0{\Sigma\vdash [A:] \dashv \Delta}
    \end{prooftree}
  \end{minipage}
\end{center}

for some concrete propositions A.
If one wants to keep an object language level distinction between the proofs of a proposition, its (trivial) type 1 enthymemes and its (non-monotonic) type 2 enthyemems one can do so by introducing a dedicated judgment for each case or a subtyping hierarchy.
However, for present purposes we don't need this separation.
Obligations and presumptions are (incomplete) proof bureaucracy and as such don't come with reduction rules.
However, we have the following $\eta$ equivalence rules:

\begin{align*}
  ([.:]-\eta&) & v &\alignleftrightarrow & \mu \alpha.\rangle v  :: \rangle \mu\_.\rangle x :: \alpha \langle::[.:]\langle .x\mu \langle \\
  ([:.]-\eta&) & E &\alignleftrightarrow & \rangle \mu \alpha.\rangle [:.] :: \rangle \alpha :: E\langle.\_\mu\langle :: E \langle .x\mu
\end{align*}


We call the language obtained in this way the \emph{Argument Calculus} (short name: \AC).\ednote{Probably too early here.}


\begin{oldpart}{Largely from old paper. This is now partially outdated.}

  \subsection{Arguments}

  Our definition of arguments is inspired by deductive argumentation \cite{besnard-review-argum} which views arguments as conditional proofs of the form $\Sigma \vdash_{\mathcal{L}} \varphi$ where the \emph{assumptions} $\Sigma$ and the \emph{conclusion} $\varphi$ are formulae of a given base logic ith consequence relation $\vdash_{\mathcal{L}}$.
  `In sequent-based variant \cite{TODO},\ednote{TODO: sequent-based argumentation citation.} they then become single conclusion sequents.
  However, we depart from this approach in two ways:
  Firstly, unlike deductive argumentation, we do not wish to abstract away from the interior structure of the arguments.
  Hence our approach is to define \emph{arguments} as ``open'' terms and \emph{counterarguments} as ``open'' environments.
  Secondly, we want to avoid overloading object-language-level variables whose main functionality is hypothetical reasoning.
  Unlike premises in conditonal proofs, argument assumptions are not merely hypothetical but closer to (defeasible) knowledge items.
  Hence their main route of elimination is not discharge but defeat.

  This introduces two distinctions from the standard deductive argumentation approach: firstly, in our setting arguments only depend on the assumptions which they use.
  In particular, irrespective of the antecedent (succedent) it is not possible to attack an argument on an assumption that does not appear free in it.
  Secondly, the same sequent represents many different arguments.
  Instead our way of identifying arguments will function via defining a normal form.
  The basic idea is that an argument is an \AC term where every subargument exposes a corresponding obligation which can be the target of debate operations such as attack or support.
  This exposure is \emph{bureaucratic} in nature in that it itself has no logical content.

  \begin{definition}[Bureaucratic Expressions]
    An \AC expression is a \emph{bureaucratic obligation} if it belongs to the following subgrammar:
    
    \begin{grammar}[Bureaucratic Obligation Expressions][][gram:bureau]
      \firstcasesubtil{$b$}{[:\sigma]}{}
      \otherform{[\sigma:]}{}
      \otherform{\mu \sigma:\alpha.\rangle b :\sigma: E \langle}{}
      \otherform{\mu \sigma:\alpha.\rangle v :\sigma: b \langle}{}
      \otherform{\rangle b :\sigma: E \langle.x : \sigma \mu}{}
      \otherform{\rangle v :\sigma: b \langle. x : \sigma \mu}{}
    \end{grammar}
  \end{definition}

  Bureaucratic obligation expressions make no logical progress but instead manage the control flow around obligations.

  \begin{example}[Bureaucratic Obligation]
    Consider the expression $\mu \alpha:A \rightarrow B.\rangle \mu\_.\rangle \lambda x. [:B] :: \alpha \langle :A\rightarrow B: [A \rightarrow B:]\langle$. Via one application of $\mu <$ it reduces to $\mu \alpha:A.\rangle \lambda x.[:B]:: \alpha \langle$ which is $\eta$-equivalent to $\lambda x.[:B]$. At the same time it exposes an obligation $[:A \rightarrow B]$ which could be filled in - for example - by a continuation that rerouts execution to use some alternative proof: $\rangle v' :A\rightarrow B: \alpha\langle.\_\mu$. In other words, we can think of an arbitrary proof term as a sceleton of obligations, fleshed out by an arbitrary number of proofs (refutations) for each sub-obligation.
  \end{example}

  \begin{definition}[Bureaucratic Obligation Redexes]
    Let $v,E$ be terms of the shapes  $\mu \alpha.\rangle \mu \_.\rangle v::\alpha\langle :: [.:] \langle$ and $ \rangle [:.] :: \rangle x:: E\langle .\_\mu\langle .x\mu$, respectively. A  \emph{bureaucratic obligation redex} is one of the following expressions:
    \[
      \mu \sigma:\alpha.\rangle  v :\sigma: E' \langle
    \]
    \[
      \rangle v' :\sigma: E \langle.x : \sigma \mu
    \]
    \[
      \rangle v :\sigma: [\sigma:] \langle :\sigma: E' \langle.x: \sigma \mu
    \]
    \[
      \mu \sigma:\alpha.\rangle v' :\sigma: E \langle
    \]
    In each case, the bound command is a default redex; hence the whole surrounding $\mu$- or $\mu'$-expression is also a redex, and moreover bureaucratic.
  \end{definition}

  \begin{definition}[obligation-$\eta$ long form]
    An expression is in \emph{obligation-$\eta$ long form} iff obligation-$\eta$-expanding any of its subterms would introduce a bureaucratic obligation redex.
  \end{definition}

  With these ingredients in place we are finally ready to define our notion of argument:

  \begin{definition}[Argument]
    Given a proposition $\sigma$ the \emph{arguments for $\sigma$} are the intuitionistic, obligation-$\eta$-long $\AC$-expressions $v:\sigma$. The \emph{arguments against $\sigma$} (or counterarguments to $\sigma$) are the dual-intuitionistic obligation-$\eta$-long $AC$ expressions $\sigma:E$. An \emph{argument} is either an argument for or against some type $t'$.
  \end{definition}

  \ednote{MR: motivate restrictions to intuitionistic/dual-instuitionistic fragments: control flow, non-interaction with debate operations.}
\end{oldpart}
\end{document}
%%% Local Variables:
%%% mode: LaTeX
%%% TeX-master: t
%%% End:
